---
title: "Reproducible biometeorological research with R (an example)"
author: "Ioannis Charalampopoulos"
date: "22/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE
)
```


# Introduction
The following is an absolutely hypothetical research scenario.

## Scope
We want to investigate the thermal comfort over an area from Toulouse (France) to Andorra Principality.

## Data, method and output
The data will be carried out from NASAPOWER data repository. The selection will be made by the definition of geographical coordinates (LAT/LON).

We will calculate the THI and Wind Chill indices for a ten year period (1995-2005) in a daily basis with rBiometeo package.
The output will be a rendered as a .pdf file containing the text and the code in grey boxes,a violin-plot in .png format and and time-series facet plot in .jpeg format both of them in specific size (width/ height).
Also, the data table will saved in a .csv file. Of, course the user can run the code directly by this .Rmd file to reproduce all the research workflow.

# Data acquisition 

We use the nasapower package to download air temperature (average, max, min) and Relative Humidity (RH) at 2m, and Wind Speed (WS) at 2 and 10 meters.


```{r data acquisition, cache=TRUE}
#Load package
library(nasapower)

# Retrieve data from NASAPOWER
my_data <- get_power(
  community = "AG",
  lonlat = c(0.83206,42.5431,2.08450,43.8485),
  pars = c("T2M","T2M_MAX","T2M_MIN","WS2M","WS10M", "RH2M"),
  dates = c("1995-01-01", "2005-12-31"),
  temporal_average = "DAILY"
)
```


# Data handling

```{r data handling, cache=TRUE}
# Load package
library(dplyr)

# Convert the downloaded data to table
my_data_tibble <- as_tibble(my_data) 

# Create a column with seasons
my_data_tb <- my_data_tibble %>%
  mutate(Season =
           ifelse(MM %in% c(1, 2, 12), "Winter",
                  ifelse(
                    MM %in% c(3, 4, 5), "Spring",
                    ifelse(MM %in% c(6, 7, 8), "Summer",
                           ifelse(MM %in% c(9, 10, 11), "Autumn", "Error"))
                  )))
```



# Data analysis

```{r Data analysis}
# Load packages
library(rBiometeo)
library(dplyr)

# Add a column with THI index
my_data_tb  <-  my_data_tb %>%
  mutate(THI = rBiometeo::THI(T2M, RH2M))

# Add a column with Wind chill index
my_data_tb  <-  my_data_tb %>%
  mutate(Windchill = rBiometeo::windchill(T2M, WS2M))

my_data_tb$YEAR <- as.integer(my_data_tb$YEAR)


# Inspect the first rows of the new dataset
head(my_data_tb)
```

# Data dissemination (results visualisation)

The results of the indices' calculations will be saved as an .csv file and will be used for two graphs (violin-plot and time series plot with facets)

```{r, data dissemination, echo=TRUE, warning=FALSE, message=FALSE, cache=TRUE}
# Create a violin plot for the THI index for all sampling points per season
# and save it in png format with specific size

## Load the library
library(ggplot2)


## Create the violin plot 
violin_plot <-
  ggplot(my_data_tb, aes(x = Season, y = THI, fill = Season)) +
  geom_violin()

## Illustrate the violin plot
plot(violin_plot)

## Save the last plot in a .png file with the below size
ggsave(
  "violin_plot.png",
  width = 40,
  height = 20,
  units = "cm"
)

## Create smoothed line graphs of the Windchill index for all the sampling 
## points per season
smooth_plot <- ggplot(my_data_tb, aes(x = YEAR, y = Windchill)) +
  geom_smooth() +
  facet_wrap(. ~ Season) +
  scale_x_continuous(breaks = c(1995, 2000, 2005)) +
  theme(axis.title.x = element_text(color = "black", size = 10, face = "bold"))

plot(smooth_plot)

## Save the last plot in a .jpeg file with the below size and resolution
ggsave(
  "time_plot.jpeg",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)


## Save the data table in a .csv file called indices.csv
write.csv(my_data_tb, file = "indices.csv")
```



